\textbf{Multi-pitch estimation based on mathematical models}\\A pitch estimation framework and harmonic sieve algorithm

### 2021-03-31 paper.tex
latest verion

Single pitch estimation can be accurately performed using simple mathematical models. Multi pitch estimation is much harder as harmonic overtones are difficult to differentiate from the fundamental notes. Because of this, most recent methods use machine learning or artificial intelligence. These methods are not well suited for real-time transcription. Also, no actual (mathematical) understanding of the problem is gained. Because of this, this paper will focus on mathematically solving this problem. Even if the methods researched in this paper might not outperform ML/AI solutions by themselves, they could be used to filter the input data for the ML/AI approaches. This might in turn improve the results of the AI/ML methods significantly.

@@

\section{Pitch estimation}
Automatic music transcription can be separated into two main problems; onset detection and pitch estimation~\cite{survey2}. In this paper, we will focus on pitch estimation. Because we always output the set of played notes for the last analysed frame, we do not require onset information. However, this information can improve the results of pitch estimation. When the duration of a note is known, we choose to only use the samples from within a note for analysis. Also, onset/offset information can be used to more effectively filter transients.

@@

@@

A similar algorithm can work for the polyphonic case. As the number of played notes in unknown, we have to determine of a note is a fundamental or harmonic. We cannot just set a threshold for the number of overtones a fundamental should have, because the number of detected overtones is very inconsistent.

To solve this, we present a harmonic sieve algorithm. Instead of directly looking for fundamentals, we sieve all guaranteed harmonics from the peaks.

@@

\section{Future work}  \label{sec:future}
How to make results more consistent (onset detection, transient filtering). How to make the project more real-time (threading rolling transforms).

>>>>

Single pitch estimation can be accurately performed using simple mathematical models~\cite{mono}. Multi pitch estimation is much harder as harmonic overtones are difficult to differentiate from the fundamental notes~\cite{oud}. Because of this, most recent methods use machine learning or artificial intelligence. These methods are not well suited for real-time transcription. Also, no actual (mathematical) understanding of the problem is gained. Because of this, this paper will focus on mathematically solving this problem. Even if the methods researched in this paper might not outperform ML/AI solutions by themselves, they could be used to filter the input data for the ML/AI approaches. This might in turn improve the results of the AI/ML methods significantly.

@@

\section{Pitch estimation}
Automatic music transcription can be separated into two main problems; onset detection and pitch estimation~\cite{survey2}. In this paper, we will focus on pitch estimation. Because we always output the set of played notes for the last analysed frame, we do not require onset information. However, this information can improve the results of pitch estimation. For example, when a onset is detected, we can discard all samples from the transient.

@@

There are many parameters which change the behaviour of the overall system. For this paper, we choose some parameters which seemed perform well, however, these parameters can likely be optimized. All parameters can be found in \texttt{config.h} along with comments explaining them.

By default, the program assumes an $A_4$ of 440 Hz. This can also be edited in \texttt{config.h}. Only 12 tone equal temperament (12-TET) tuning is currently supported. It could easily be changed to n-TET tuning, but non TET tuning is not easily supported.

@@

Due to small errors, the overtones are not exactly integer multiples of the fundamental. As the distance between notes increases with frequency, we cannot set a constant threshold of error in Hz. Instead, we calculate the error in cents and choose a threshold for this error (configurable in \texttt{config.h}). This also allows for equal error tolerance above and below the note, which percentage based methods would not.

Polyphonic note detection can use a algorithm similar to the monophonic algorithm. As the number of played notes in unknown in the polyphonic case, we have to determine if a note is a fundamental or harmonic. We cannot just set a threshold for the number of overtones a fundamental should have, because the number of detected overtones is very inconsistent. To solve this, we present a harmonic sieve algorithm. Instead of directly looking for fundamentals, we try to sieve away all harmonics until we are left with fundamentals. We start at the lowest peak end mark all its possible overtones. If there are still unmarked peaks left, they are likely from another note. We can repeat this process until there are no peaks left.

This algorithm has some problems in its current form. Solving these is outside of the scope of this paper and will be the topic of the follow-up thesis, but we will address the problems here. The first big problem is transients.

@@

\section{Conclusions}
adsf

\section{Future work}  \label{sec:future}
How to make results more consistent (onset detection, transient filtering). How to make the project more real-time (threading rolling transforms). Also deliberately detuning strings for better results.
